---
title: "Treasurer's End of Year Report"
author: "David Barron"
date: "30 June 2023"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)

opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = '')

inline_hook <- function(x){
  
  format(x, nsmall = 2, scientific = FALSE, big.mark = ",")
  
  
}

knit_hooks$set(inline = inline_hook)

```



```{r pastdata, echo=FALSE, message=FALSE}

library(readxl)
library(tidyverse)

dta <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2022-23\\TASO 2022-23.xlsm",
                 sheet = "Transactions",
                 skip = 4,
                 col_names = c("Date", "Ref", "Details", "Expenditure", "Income",
                               "Balance", "Income Category", "Expenditure Category",
                               "Funding Category", "I/E", "Amount", "I/E Category"
                               ),
                 col_types = c("date", rep("guess", 11)))

nrows <- nrow(dta)
Bank_Balance <- as.numeric(dta[nrows, "Balance"])

dta.sum <- dta %>%
          mutate(Date = lubridate::ymd(Date),
                 Month = lubridate::month(Date)) %>%
        group_by(Month) %>%
        summarise(
         Expend = sum(Expenditure, na.rm = TRUE),
         Inc = sum(Income, na.rm = TRUE)
        ) %>%
        mutate(Month = case_when(
                          Month == 7 ~ 1,
                          Month == 8 ~ 2,
                          Month == 9 ~ 3,
                          Month == 10 ~ 4,
                          Month == 11 ~ 5,
                          Month == 12 ~ 6, 
                          Month == 1 ~ 7,
                          Month == 2 ~ 8,
                          Month == 3 ~ 9,
                          Month == 4 ~ 10,
                          Month == 5 ~ 11,
                          Month == 6 ~ 12)) %>%
       arrange(Month)

# Reorder rows

  
  
  #dta.sum$Month <- as.numeric(row.names(dta.sum))

dta.sum <-  dta.sum %>%
    mutate(
          Income = cumsum(Inc),
          Expenditure = cumsum(Expend)
        ) %>%
        select(Month, Income, Expenditure)


dta.long <- dta.sum %>% pivot_longer(
              cols = Income:Expenditure,
              names_to = "Type",
              values_to = "Amount"
              ) 
  

          

```


```{r, budget}

budget2223 <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2022-23\\TASO 2022-23.xlsm",
                 sheet = "Budget v Actuals",
                 range = "C3:E22")

actuals2223 <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2022-23\\TASO 2022-23.xlsm",
                 sheet = "Budget v Actuals",
                 range = "G3:I22")

budgetcats <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2022-23\\TASO 2022-23.xlsm",
                 sheet = "Budget v Actuals",
                 range = "A4:A22",
                 col_names = c("Category"))

budget2324 <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2023-24\\TASO 2023-24.xlsm",
                 sheet = "Budget v Actuals",
                 range = "C3:E22")

actuals2324 <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2023-24\\TASO 2023-24.xlsm",
                 sheet = "Budget v Actuals",
                 range = "G3:I22")


budgetcats24 <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2023-24\\TASO 2023-24.xlsm",
                 sheet = "Budget v Actuals",
                 range = "A4:A21",
                 col_names = c("Category"))


```


```{r currentdata,  eval=FALSE}

dta.current <- read_xlsx("C:\\Users\\dnbar\\Dropbox\\TASO\\Accounts 2023-24\\TASO 2023-24.xlsm",
                 sheet = "Transactions",
                 skip = 4,
                 col_names = c("Date", "Ref", "Details", "Expenditure", "Income",
                               "Balance", "Income Category", "Expenditure Category",
                               "Funding Category", "I/E", "Amount", "I/E Category"
                               ),
                 col_types = c("date", rep("guess", 11)))

nrows <- nrow(dta.current)
Bank_Balance <- as.numeric(dta.current[nrows, "Balance"])

dta.current.sum <- dta.current %>%
          mutate(Date = lubridate::ymd(Date),
                 Month = lubridate::month(Date)) %>%
        group_by(Month) %>%
        summarise(
         Expend = sum(Expenditure, na.rm = TRUE),
         Inc = sum(Income, na.rm = TRUE)
        ) %>%
        mutate(Month = case_when(
                          Month == 7 ~ 1,
                          Month == 8 ~ 2,
                          Month == 9 ~ 3,
                          Month == 10 ~ 4,
                          Month == 11 ~ 5,
                          Month == 12 ~ 6, 
                          Month == 1 ~ 7,
                          Month == 2 ~ 8,
                          Month == 3 ~ 9,
                          Month == 4 ~ 10,
                          Month == 5 ~ 11,
                          Month == 6 ~ 12)) %>%
       arrange(Month)

# Reorder rows

  
  
  #dta.sum$Month <- as.numeric(row.names(dta.sum))

dta.current.sum <-  dta.current.sum %>%
    mutate(
          Income = cumsum(Inc),
          Expenditure = cumsum(Expend)
        ) %>%
        select(Month, Income, Expenditure)


dta.current.long <- dta.current.sum %>% pivot_longer(
              cols = Income:Expenditure,
              names_to = "Type",
              values_to = "Amount"
              ) 


```



```{r plot, echo=FALSE}
num.months <- dim(dta.sum)[1]

all.months <- c(month.abb[7:12], month.abb[1:6])
labels = all.months[1:num.months]

dta.long  %>%
  ggplot(aes(x = Month, y = Amount, colour = Type)) + 
          geom_line() + scale_x_continuous(
            breaks = 1:num.months,
            label =  labels
          ) + 
  scale_y_continuous(label = scales::label_dollar(prefix="£")) +
  theme_light() +
  theme(legend.position=c(0.2, 0.85), legend.title = element_blank()) 
          


```

This report provides a summary of income and expenditure on a striclty cash basis for the TASO financial year 2022-23.  No account is taken of payments made or received that relate to activities in a different financial year (although such adjustments are of course made in our annual accounts).

---

## Budget

- Budgeted income: £`r as.numeric(budget2223[19, 1])`

- Budgeted expenditure: £`r as.numeric(budget2223[19, 2])`

- Budgeted deficit: £`r -(as.numeric(budget2223[19, 3]))`




```{r showbudget}

budop <- bind_cols(budgetcats, budget2223, actuals2223,
                   .name_repair = "minimal")

budop %>% kbl() %>%
  kable_classic(full_width = FALSE, "striped") %>%
  column_spec(4, border_right = TRUE) %>%
  add_header_above(c(" ","Budget" = 3, "Actuals" = 3))  %>%
  as_image(width = 6)

```


## Income and Expenditure

- Total income in 2022-23: £`r as.numeric(dta.sum[num.months, 2])`

- Total expenditure is 2022-23: £`r as.numeric(dta.sum[num.months, 3])`

- **Financial year end deficit is £`r -as.numeric(dta.sum[num.months,2] - dta.sum[num.months,3])`**. 

- Bank balance on 30 June 2023: £`r Bank_Balance`.

- Number of subscriptions paid by 30 June: 
  `r format((actuals2223[2,3] / 45), nsmall = 0)`
  
## Comments

-  Subscriptions received to date are about £4k lower than anticipated.

-  Charitable grants still to be paid come to £`r 14100 - 12360`.

-  The invoice from Magdalen for June's lecture has not yet been received.

-  Deposits have been paid to Rewley House for 2023-24's study days. 
